---
title: "Results Section: Public Metadata"
output:
  pdf_document: default
---
```{r}
library(staphopia)
library(ggplot2)
library(reshape2)
USE_DEV = TRUE
```

## Aggregating Data For Public Samples
First we'll get all publicly available *S. aureus* samples.

```{r}
ps <- get_public_samples()
```

## Variation From *S. aureus* N315
In Staphopia all samples had variants (SNPs and InDels) called using *S. aureus* N315 as the reference genome. In this section we'll visualize the total number of variants each sample has. This will give us an idea of the sequenced genitic diversity with respect to N315.

### Gather Variant Counts
We will use `get_variant_counts()` to get the variant counts for each sample. We will also order the counts by the total.

```{r}
variant_counts <- get_variant_counts(ps$sample_id)
variant_counts <- variant_counts[order(total),]
```

### Summary of Variant Counts
#### Total Variants (SNPs and InDels)
```{r}
summary(variant_counts$total)
```

#### SNPs
```{r}
summary(variant_counts$snp_count)
```

#### InDels
```{r}
summary(variant_counts$indel_count)
```

### Visualizing Variant Counts
#### Total Variants (SNPs and InDels)
```{r fig.width=12, fig.asp=0.4}
p <- ggplot(data=variant_counts, aes(x=seq(1,nrow(variant_counts)), y=total)) +
    xlab("Sample Count") +
    ylab("Variant Count)") +
    geom_bar(stat='identity') +
    scale_x_continuous(breaks = seq(0, nrow(variant_counts), by = 5000)) +
    scale_y_continuous(breaks = seq(0, max(variant_counts$total), by=10000), labels = scales::comma) +
    theme_bw() +
    theme(axis.text=element_text(size=12),
          axis.title=element_text(size=14,face="bold"))
p
```

#### SNPs Only
```{r fig.width=12, fig.asp=0.4}
p <- ggplot(data=variant_counts, aes(x=seq(1,nrow(variant_counts)), y=snp_count)) +
    xlab("Sample Count") +
    ylab("SNP Count") +
    geom_bar(stat='identity') +
    scale_x_continuous(breaks = seq(0, nrow(variant_counts), by = 5000)) +
    scale_y_continuous(breaks = seq(0, max(variant_counts$snp_count), by=10000), labels = scales::comma) +
    theme_bw() +
    theme(axis.text=element_text(size=12),
          axis.title=element_text(size=14,face="bold"))
p

# Output plot to PDF and PNG
staphopia::write_plot(p, paste0(getwd(), '/images/figure-x-snp-accumulation'))
```

#### InDels Only
```{r fig.width=12, fig.asp=0.4}
p <- ggplot(data=variant_counts, aes(x=seq(1,nrow(variant_counts)), y=indel_count)) +
    xlab("Sample Count") +
    ylab("Indel Count") +
    geom_bar(stat='identity') +
    scale_x_continuous(breaks = seq(0, nrow(variant_counts), by = 5000)) +
    scale_y_continuous(breaks = seq(0, max(variant_counts$indel_count), by=500), labels = scales::comma) +
    theme_bw() +
    theme(axis.text=element_text(size=12),
          axis.title=element_text(size=14,face="bold"))
p
```
