---
title: "Results Section: Pipeline Design and Processing 43,000+ genomes"
output:
  pdf_document: default
---

In this notebook will be generating statistics and plots related to processing 43,000+ 
genomes on Seven Bridges Cancer Genomics Cloud (CGC) platform.

### Load Up Packages
```{r}
library(staphopia)
library(ggplot2)
library(dplyr)
```

### Read In The Data
```{r}
results <- read.table("../data/cgc-runs.txt", header = TRUE, sep = "\t")
colnames(results)
```

This leaves use with 9 columns:

1. name: Name of the job
2. status: Job's status
3. project: CGC project job was executed from.
4. app: CGC app used to execute the job.
5. created_by: User who submitted the job.
6. total_time: Total amount of time (in minutes) a job was queued and run
7. run_time: Total amount of time (in minutes) a job took to complete
8. queue_time: Total amount of time (in minutes) a job was queued
9. price: Total cost of the run

### Clean Up The Data
Before we generate statistics and plots, we need to clean the data. There are jobs where the *run_time* 
and *price* were not properly reported from CGC. We will filter samples where the *run_time* is 0.

```{r}
results_clean <- results[results$run_time > 0, ]
nrow(results) - nrow(results_clean)
```

### Job Summary
#### Run Time Summary
```{r}
summary(results_clean$run_time)
```

#### Number of Jobs With > 120 Minute Runtime
```{r}
nrow(results_clean[results_clean$run_time > 120, ])
```

#### Summary of Jobs With Run Time Between 10 and 120 Minutes
```{r}
summary(results_clean[between(results_clean$run_time, 10, 120), ]$run_time)
```

### Plots
#### Run Time (Complete)
```{r fig.width=12, fig.asp=0.4}
p <- ggplot(data=results_clean, aes(run_time)) +
    xlab("Run Time (In Minutes)") +
    ylab("Count") +
    geom_histogram(bins=100) +
    theme_bw() +
    theme(axis.text=element_text(size=12),
          axis.title=element_text(size=14,face="bold"))
p
```

#### Pipeline Run Time (Between 10-120 Minutes)
```{r Figure 2 Pipeline Run Time, fig.width=12, fig.asp=0.4}
p <- ggplot(data=results_clean[between(results_clean$run_time, 10, 120),],
            aes(run_time)) +
    xlab("Run Time (In Minutes)") +
    ylab("Count") +
    geom_histogram(bins=100) +
    theme_bw() +
    theme(axis.text=element_text(size=12),
          axis.title=element_text(size=14,face="bold"))
p

# Output plot to PDF and PNG
staphopia::write_plot(p, paste0(getwd(), '/../figures/supplementary-figure-02-pipeline-run-time'))
```


## Session Info
```{r}
sessionInfo()
```
