---
title: "Results Section: Public Sequencing Metrics"
output:
  pdf_document: default
---
```{r}
library(staphopia)
library(ggplot2)
library(reshape2)
```

## Aggregating Data For Public Samples
First we'll get all publicly available *S. aureus* samples.

```{r}
ps <- get_public_samples()
```

We will also get information pertaining to submissions and ranks by year.
```{r}
submissions <- get_submission_by_year(all = TRUE)
ranks <- get_rank_by_year()
```

We now have `r toString(nrow(ps))` samples to work with. Next we will acquire
metadata, sequencing stats and assembly stats associated with each sample.

```{r}
metrics <- merge(
    ps,
    merge(
        get_assembly_stats(ps$sample_id),
        merge(
            get_metadata(ps$sample_id),
            get_sequence_quality(ps$sample_id, stage='cleanup'),
            by='sample_id'
        ),
        by='sample_id'
    ),
    by='sample_id'
)
```

We are now going to add two columns `rank_name` and `year`.
```{r}
metrics$year <- sapply(
    metrics$first_public,
    function(x) {
        strsplit(x, "-")[[1]][1]
    }
)

metrics$rank_name <- ifelse(
    metrics$rank.x == 3,
    'Gold',
    ifelse(
        metrics$rank.x == 2,
        'Silver',
        'Bronze'
    )
)
```

## Visualizing Metrics
The following sections will be plots to visualize relationships in the data.

### By Year Plots
#### Submissions Per Year
```{r fig.width=12, fig.asp=0.4}
p <- ggplot(data=submissions, aes(x=year, y=count)) +
    xlab("Year") +
    ylab("Count") +
    geom_bar(stat='identity') +
    geom_text(aes(label=count), vjust = -0.5) +
    scale_x_continuous(breaks = round(seq(min(submissions$year), max(submissions$year), by = 1),1)) +
    theme_bw() +
    theme(axis.text=element_text(size=12),
          axis.title=element_text(size=14,face="bold"))
p
```

#### Overall Submissions
```{r Figure 1, fig.asp=0.4, fig.width=12}
p <- ggplot(data=submissions, aes(x=year, y=overall)) +
    xlab("Year") +
    ylab("Cumulative Count") +
    geom_bar(stat='identity') +
    geom_text(aes(label=overall), vjust = -0.5) +
    scale_x_continuous(breaks = round(seq(min(submissions$year), max(submissions$year), by = 1),1)) +
    theme_bw() +
    theme(axis.text=element_text(size=12),
          axis.title=element_text(size=14,face="bold"))
p

# Output plot to PDF and PNG
staphopia::write_plot(p, paste0(getwd(), '/../figures/figure-01-submissions-per-year'))
```

#### Submission Ranks
```{r Figure 4 BSG, fig.width=12, fig.asp=0.5}
melted <- melt(ranks, id=c('year'),
               measure.vars = c('bronze', 'silver', 'gold'))
melted$title <- ifelse(melted$variable == 'gold', 'Gold', 
                       ifelse(melted$variable == 'silver', 'Silver', 'Bronze'))
melted$rank <- ifelse(melted$variable == 'gold', 3, 
                      ifelse(melted$variable == 'silver', 2, 1))
p <- ggplot(data=melted, aes(x=year, y=value, fill=title, group=rank, label=title)) +
    xlab("Year") +
    ylab("Count") +
    geom_bar(stat='identity', position='dodge') +
    geom_text(aes(label=value), vjust = -0.5, position = position_dodge(.9)) +
    scale_fill_manual(values=c("#CD7F32", "#D4AF37", "#C0C0C0")) +
    scale_x_continuous(breaks = round(seq(min(ranks$year), max(ranks$year), by = 1),1)) +
    theme_bw() +
    theme(axis.text=element_text(size=12),
          axis.title=element_text(size=14,face="bold"),
          legend.title = element_blank())
    

p

# Output plot to PDF and PNG
staphopia::write_plot(p, paste0(getwd(), '/../figures/figure-04-rank-per-year'))
```

#### Assembly Size
```{r fig.width=12, fig.asp=0.5}
p <- ggplot(metrics, aes(x = year, y = total_contig_length)) +
    geom_boxplot()
p
```

#### Total Contigs (smaller is better)
```{r fig.width=12, fig.asp=0.5}
p <- ggplot(metrics, aes(x = year, y = total_contig)) +
    geom_boxplot()
p
```

#### N50
```{r fig.width=12, fig.asp=0.5}
p <- ggplot(metrics, aes(x = year, y = n50_contig_length)) +
    geom_boxplot()
p
```

#### Mean Contig Length
```{r fig.width=12, fig.asp=0.5}
p <- ggplot(metrics, aes(x = year, y = mean_contig_length)) +
    geom_boxplot()
p
```

#### Max Contig Length
```{r fig.width=12, fig.asp=0.5}
p <- ggplot(metrics, aes(x = year, y = max_contig_length)) +
    geom_boxplot()
p
```

#### Mean Read Length
```{r fig.width=12, fig.asp=0.5}
p <- ggplot(metrics, aes(x = year, y = read_mean)) +
    geom_boxplot()
p
```

#### Mean Per-Read Quality Score
```{r fig.width=12, fig.asp=0.5}
p <- ggplot(metrics, aes(x = year, y = qual_mean)) +
    geom_boxplot()
p
```

#### Assembly Size Grouped By Rank
```{r fig.width=12, fig.asp=0.5}
p <- ggplot(metrics, aes(x = year, y = total_contig_length,
                         fill=rank_name, label=rank_name)) +
    geom_boxplot()
p
```

#### Total Contigs Grouped By Rank
```{r fig.width=12, fig.asp=0.5}
p <- ggplot(metrics, aes(x = year, y = total_contig,
                         fill=rank_name, label=rank_name)) +
    geom_boxplot()
p
```

#### N50 Grouped By Rank
```{r fig.width=12, fig.asp=0.5}
p <- ggplot(metrics, aes(x = year, y = n50_contig_length,
                         fill=rank_name, label=rank_name)) +
    geom_boxplot()
p
```

#### Mean Contig Length Grouped By Rank
```{r fig.width=12, fig.asp=0.5}
p <- ggplot(metrics, aes(x = year, y = mean_contig_length,
                         fill=rank_name, label=rank_name)) +
    geom_boxplot()
p
```

#### Max Contig Length Grouped By Rank
```{r fig.width=12, fig.asp=0.5}
p <- ggplot(metrics, aes(x = year, y = max_contig_length,
                         fill=rank_name, label=rank_name)) +
    geom_boxplot()
p
```

#### Mean Read Length Grouped By Rank
```{r fig.width=12, fig.asp=0.5}
p <- ggplot(metrics, aes(x = year, y = read_mean,
                         fill=rank_name, label=rank_name)) +
    geom_boxplot()
p
```

#### Mean Per-Read Quality Score Grouped By Rank
```{r fig.width=12, fig.asp=0.5}
p <- ggplot(metrics, aes(x = year, y = qual_mean,
                         fill=rank_name, label=rank_name)) +
    geom_boxplot()
p
```

### By Rank Plots
#### Assembly Size
```{r fig.width=12, fig.asp=0.5}
p <- ggplot(metrics, aes(x = rank_name, y = total_contig_length)) +
    geom_boxplot()
p
```

#### Total Contigs (smaller is better)
```{r fig.width=12, fig.asp=0.5}
p <- ggplot(metrics, aes(x = rank_name, y = total_contig)) +
    geom_boxplot()
p
```

#### N50
```{r fig.width=12, fig.asp=0.5}
p <- ggplot(metrics, aes(x = rank_name, y = n50_contig_length)) +
    geom_boxplot()
p
```

#### Mean Contig Length
```{r fig.width=12, fig.asp=0.5}
p <- ggplot(metrics, aes(x = rank_name, y = mean_contig_length)) +
    geom_boxplot()
p
```

#### Max Contig Length
```{r fig.width=12, fig.asp=0.5}
p <- ggplot(metrics, aes(x = rank_name, y = max_contig_length)) +
    geom_boxplot()
p
```

#### Mean Read Length
```{r fig.width=12, fig.asp=0.5}
p <- ggplot(metrics, aes(x = rank_name, y = read_mean)) +
    geom_boxplot()
p
```

#### Mean Per-Read Quality Score
```{r fig.width=12, fig.asp=0.5}
p <- ggplot(metrics, aes(x = rank_name, y = qual_mean)) +
    geom_boxplot()
p
```

#### Coverage
```{r fig.width=12, fig.asp=0.5}
p <- ggplot(metrics, aes(x = rank_name, y = coverage)) +
    geom_boxplot()
p
```

## Bronze Data
#### Coverage By Quality
```{r fig.width=12, fig.asp=0.5}
p <- ggplot(metrics[metrics$rank.x == 1,], aes(x = coverage, y = qual_mean)) +
    geom_point()
p
```

#### Coverage By Read Length
```{r fig.width=12, fig.asp=0.5}
p <- ggplot(metrics[metrics$rank.x == 1,], aes(x = coverage, y = read_mean)) +
    geom_point()
p
```


## Session Info
```{r}
sessionInfo()
```
