---
title: "Results Section: Public Genetic Diversity"
output:
  pdf_document: default
---
```{r}
library(staphopia)
library(ggplot2)
library(reshape2)
library(scales)
```

In this section we will look into genetic diversity that has been sequenced in *Staphylococcus aureus*. In order to do so, we'll use variant counts, cgMLST and MLST as measures of diversity.


## Aggregating Data For Public Samples
First we'll get all publicly available *S. aureus* samples.

```{r}
ps <- get_public_samples()
```

## MLST
Next we will will use the MLST information has a measure of genitic diversity. In this case we are interested in the total number of unique sequence types sequenced. We'll use *get_st_by_year()* to get some basic stats about how many STs have been sequenced. We will also use *get_top_sequence_types()* to get each ST represented in the database and the total number of samples with each ST. (*Note: 5000 is just an arbitrarly large number to retreive all STs*)
```{r}
sequence_types <- get_st_by_year()
top_st <- get_top_sequence_types(5000)
colnames(sequence_types)
```

This gives us 38 columns for each year. These columns are:

1. year: The year.
2. unique: The Number of unique STs for a given year.
3. novel: Number of STs not sequenced previously.
4. assigned: Samples which a ST was determined.
5. assigned_agree: Samples in which each program that called an ST agreed in ST.
6. assigned_disagree: Samples in which programs did not each call the same ST.
7. unassigned: Samples which a ST was not determined.
8. unassigned_agree: Each program was unable to assign an ST.
9. unassigned_disagree: Samples in which no ST was determined, but each program does not agree
10. predicted_novel: Samples with a match to each Loci, but allele pattern does not exist.
11. all: Samples with an ST determined with agreement between each program.
12. partial: Samples with an ST determined with agreement between two programs.
13: ariba_blast: Samples with an ST determined with agreement between Ariba and BLAST.
14. mentalist_blast: Samples with an ST determined with agreement between MentaLiST and BLAST.
15. mentalist_ariba: Samples with an ST determined with agreement between MentaLiST and Ariba.
16. single: Samples with an ST determined by only a single program.
17. ariba: Samples with an ST determined by only Ariba.
18. mentalist: Samples with an ST determined by only MentaLiST.
19. blast: Samples with an ST determined by only BLAST.
20. count: Total number of samples in a given year.
21-38: overall_*X*: The cumulative totals of previous years for column *x*

#### Compare MLST Predictions
```{r}
mlst <- get_sequence_type(ps$sample_id)
metadata <- merge(
    ps, 
    get_metadata(ps$sample_id),
    by='sample_id'
)
metadata$year <- sapply(
    metadata$first_public,
    function(x) {
        strsplit(x, "-")[[1]][1]
    }
)

metadata$rank_name <- ifelse(
    metadata$rank == 3,
    'Gold',
    ifelse(
        metadata$rank == 2,
        'Silver',
        'Bronze'
    )
)
```


```{r}
mlst_temp <- merge(mlst, metadata[,c('sample_id', 'is_paired')], by='sample_id')
mlst_temp$is_paired <- ifelse(mlst_temp$is_paired == "", FALSE, TRUE)


mlst_temp$agreement <- paste0(
    ifelse(mlst$st == 0 | mlst$st == 0, '000',
        ifelse(mlst$mentalist == mlst$ariba & mlst$mentalist == mlst$blast, '111', 
            ifelse(mlst$mentalist == mlst$ariba, '110', 
                ifelse(mlst$mentalist == mlst$blast, '101',
                    ifelse(mlst$blast == mlst$ariba, '011',
                        ifelse(mlst$mentalist > 0, '100', 
                            ifelse(mlst$ariba > 0, '010',
                                ifelse(mlst$blast > 0, '001', '000')
                            )
                        )
                    )
                )
            )
        )
    )
)

# mentalist
# ariba
# blast
mlst_temp$agreement<- ifelse(mlst_temp$is_paired == TRUE, mlst_temp$agreement,
                             paste0(substr(mlst_temp$agreement, 1, 1), '-',
                                    substr(mlst_temp$agreement, 3, 3)))
table(mlst_temp$agreement)
platform <- metadata[,c('sample_id', 'instrument_model', 'study_accession', 'year', 'rank_name')]
mlst_temp <- merge(mlst_temp, platform, by='sample_id')
```

```{r}
table(mlst_temp[mlst_temp$agreement == '101',]$rank_name)
table(mlst_temp[mlst_temp$agreement == '011',]$rank_name)
table(mlst_temp[mlst_temp$agreement == '110',]$rank_name)
```

### PubMLST ST Counts
```{r}
st_counts <- merge(
    read.table('../data/pubmlst-counts.txt', header=TRUE, sep="\t"),
    staphopia <- top_st[top_st$st > 0,c('st', 'count')],
    by='st', all=TRUE
)
st_counts[is.na(st_counts)] <- 0
st_counts <- st_counts[st_counts$pubmlst_count > 0,]
nrow(st_counts)
nrow(st_counts[st_counts$count == 0,])
nrow(st_counts[st_counts$count >= 1,])
nrow(st_counts[st_counts$pubmlst_count == 1,])
nrow(st_counts[st_counts$pubmlst_count <= 2,])
table(st_counts[st_counts$count == 0,]$pubmlst_count)
summary(st_counts[st_counts$count > 0,]$pubmlst_count)
st_counts[st_counts$pubmlst_count > 10 & st_counts$count == 0,]
```

### Summary of MLST Diversity
#### Assignment Breakdown
```{r}
t(sequence_types[sequence_types$year == max(sequence_types$year),21:38])
```

#### Top STs
```{r}
top_st[1:10,]
```

This gives us 4 columns for each ST, in descending order based on the *count* column. In other words the most represented STs are seen first. These columns are:

1. st: The sequence type.
2. count: The number of samples with given ST.
3. percent: The percent of samples represented by given ST.
4. overall: The percent of samples represented by given ST and previous STs.

##### How many unique STs represented?
```{r}
nrow(top_st[top_st$st > 0,])
```

##### How many STs represented by a single sample?
```{r}
nrow(top_st[top_st$count == 1, ])
```


###  Visualizing MLST Diversity
The following sections will be plots to visualize relationships in the data.

#### Unique Sequence Types By Year
```{r fig.width=12, fig.asp=0.4}
p <- ggplot(data=sequence_types, aes(x=year, y=unique)) +
    xlab("Year") +
    ylab("Count") +
    geom_bar(stat='identity') +
    geom_text(aes(label=unique), vjust = -0.5) +
    scale_x_continuous(breaks = round(seq(min(sequence_types$year), 
                                          max(sequence_types$year), 
                                          by = 1),1)) +
    theme_bw() +
    theme(axis.text=element_text(size=12),
          axis.title=element_text(size=14,face="bold"))
p
```

#### Novel Sequence Types By Year
```{r fig.width=12, fig.asp=0.4}
p <- ggplot(data=sequence_types, aes(x=year, y=novel)) +
    xlab("Year") +
    ylab("Count") +
    geom_bar(stat='identity') +
    geom_text(aes(label=novel), vjust = -0.5) +
    scale_x_continuous(breaks = round(seq(min(sequence_types$year), 
                                          max(sequence_types$year), 
                                          by = 1),1)) +
    theme_bw() +
    theme(axis.text=element_text(size=12),
          axis.title=element_text(size=14,face="bold"))
p
```

#### Overall Novel Sequence Types By Year
```{r fig.width=12, fig.asp=0.4}
p <- ggplot(data=sequence_types, aes(x=year, y=overall_novel)) +
    xlab("Year") +
    ylab("Cumulative Count") +
    geom_bar(stat='identity') +
    geom_text(aes(label=overall_novel), vjust = -0.5) +
    scale_x_continuous(breaks = round(seq(min(sequence_types$year), 
                                          max(sequence_types$year), 
                                          by = 1),1)) +
    theme_bw() +
    theme(axis.text=element_text(size=12),
          axis.title=element_text(size=14,face="bold"))
p
```


#### Top 10 Sequence Types
```{r Figure 5 TopST, fig.width=12, fig.asp=0.45}
p <- ggplot(data=top_st[1:10,], aes(x=reorder(st, -count), y=count)) +
    xlab("Sequence Type") +
    ylab("Count") +
    geom_bar(stat="identity") +
    geom_text(aes(label=count), vjust = -0.5) + 
    theme_bw() +
    theme(axis.text=element_text(size=12),
          axis.title=element_text(size=14,face="bold"))
p
```

#### Total Allele Matches For Unassigned Samples
```{r fig.width=12, fig.asp=0.45}
allele_matches <- get_mlst_allele_matches(ps[ps$st == 0,]$sample_id)
df <- as.data.frame(table(allele_matches[allele_matches$matches < 7,]$matches))
colnames(df) <- c("matches", "count")

p <- ggplot(data=df, aes(x=matches, y=count)) +
    xlab("Matched Alleles") +
    ylab("Count") +
    geom_bar(stat="identity") +
    geom_text(aes(label=count), vjust = -0.5) + 
    theme_bw() +
    theme(axis.text=element_text(size=12),
          axis.title=element_text(size=14,face="bold"))
p
```

## cgMLST Patterns
Finally, we'll look at cgMLST as a measure of genetic diversity. We will use the *get_cgmlst()* function to get the cgMLST results for each Sample. This function might take a little while to retrieve all teh results.
```{r}
# USE_DEV to prevent timeout here until problem resolved
USE_DEV = TRUE

cgmlst <- get_public_cgmlst_patterns()
cgmlst$percent <- cgmlst$count / sum(cgmlst$total_samples)
cgmlst
```

This gives us two columns:

1. samples_in_pattern: The number of samples with a given cgMLST pattern.
2. count: The number patterns with a given number of samples.
3. total_samples: Number of samples represented by a row (samples_in_pattern * count)
4. percent: Percent of samples represented

For example, if samples_in_pattern is 100 and the count is 2. That means there are **2** (count=2) cgMLST patterns that are shared by **100 samples** (samples_in_count=100) each, representing a total of **200 samples** (count * samples_in_count).

### Total Number of Distinct cgMLST Patterns
```{r}
sum(cgmlst$count)
```

### How many shared cgMLST patterns?
```{r}
sum(cgmlst[cgmlst$samples_in_pattern > 1, ]$count)
```

### How many samples share a cgMLST pattern?
```{r}
sum(cgmlst[cgmlst$samples_in_pattern > 1, ]$total_samples)
```

### How many samples have a unique cgMLST pattern?
```{r}
cgmlst$percent <- cgmlst$count / sum(cgmlst$total_samples)
cgmlst[cgmlst$samples_in_pattern == 1, ]
```


## Session Info
```{r}
sessionInfo()
```
