---
title: "Results Section: Public Metadata"
output:
  pdf_document: default
---
```{r}
library(staphopia)
library(dplyr)
library(ggplot2)
library(reshape2)
```

## Aggregating Data For Public Samples
First we'll get all publicly available *S. aureus* samples.

```{r}
ps <- get_public_samples()
```

We now have `r toString(nrow(ps))` samples to work with. Next we will acquire
metadata associated with each sample.

We will also get information pertaining to submissions by year and how any publication links were made.
```{r}
submissions <- get_submission_by_year()
publication_links <- get_publication_links()
```

Next we are going to pull down any metadata associated with the public samples.
```{r}
metrics <- merge(
    ps, 
    get_metadata(ps$sample_id),
    by='sample_id'
)
```

We are now going to add two columns `rank_name` and `year`.
```{r}
metrics$year <- sapply(
    metrics$first_public,
    function(x) {
        strsplit(x, "-")[[1]][1]
    }
)

metrics$rank_name <- ifelse(
    metrics$rank.x == 3,
    'Gold',
    ifelse(
        metrics$rank.x == 2,
        'Silver',
        'Bronze'
    )
)
```

### Publication Information

#### Summary
Here are details looking at total submissions and their publication status.
```{r}
t(submissions[submissions$year == max(submissions$year),])
```

Here is information on how publication links were made.
```{r}
t(publication_links)
```

There are 6 rows and their names are as follows:

1. elink: Number samples linked to a PubMed ID identified from eLink
2. text: Number samples linked to a PubMed ID identified from text mining (not through eLink)
3. elink_pmid: Number of PubMed IDs identified from eLink
4. text_pmid: Number of PubMed IDs identified from text mining (not through eLink)
5. total: Total number of samples associated with a PubMed ID
6. total_pmid: Total number of PubMed IDs associated with published samples

##### Percent of Samples Published
```{r}
stats <- submissions[submissions$year == max(submissions$year),]
stats$overall_published / stats$overall * 100
```

#### Published vs Unpublished Submissions Per Year
```{r fig.width=12, fig.asp=0.5}
melted <- melt(submissions, id=c('year'),
               measure.vars = c('published', 'unpublished'))
melted$title <- ifelse(melted$variable == 'published', 'Published', 'Unpublished')
p <- ggplot(data=melted, aes(x=year, y=value, fill=title)) +
    xlab("Year") +
    ylab("Count") +
    geom_bar(stat='identity', position='dodge') +
    geom_text(aes(label=value), vjust = -0.5, position = position_dodge(.9)) +
    scale_fill_manual(values=c("#2ca25f", "#5ab4ac")) +
    scale_x_continuous(breaks = round(
        seq(min(submissions$year), max(submissions$year), by = 1), 1
    )) +
    theme_bw() +
    theme(axis.text=element_text(size=12),
          axis.title=element_text(size=14,face="bold"),
          legend.title = element_blank())
    

p
```

#### Overall Published vs Unpublished Submissions
```{r Figure 8 Pubs, fig.width=12, fig.asp=0.5}
melted <- melt(submissions, id=c('year'),
               measure.vars = c('overall_published', 'overall_unpublished'))
melted$title <- ifelse(melted$variable == 'overall_published', 'Published', 'Unpublished')
p <- ggplot(data=melted, aes(x=year, y=value, fill=title)) +
    xlab("Year") +
    ylab("Cumulative Count") +
    geom_bar(stat='identity', position='dodge') +
    geom_text(aes(label=value), vjust = -0.5, position = position_dodge(.9)) +
    scale_fill_manual(values=c("#2ca25f", "#5ab4ac")) +
    scale_x_continuous(breaks = round(
        seq(min(submissions$year), max(submissions$year), by = 1), 1
    )) +
    theme_bw() +
    theme(axis.text=element_text(size=12),
          axis.title=element_text(size=14,face="bold"),
          legend.title = element_blank())
    

p

# Output plot to PDF and PNG
staphopia::write_plot(p, paste0(getwd(), '/../figures/figure-08-published-per-year'))
```


### Metadata Information

#### Number of Samples With A Collection Date
```{r}
has_collection_date <- nrow(metrics[metrics$collection_date != "",])
paste0(has_collection_date," (", has_collection_date / nrow(metrics) * 100, " %)")
```

#### Number of Samples With A Location Information
```{r}
has_location <- nrow(metrics[metrics$location != "unknown/missing",])
paste0(has_location," (", has_location / nrow(metrics) * 100, " %)")
```

#### Number of Locations
```{r}
nrow(as.data.frame(table(metrics[metrics$location != "unknown/missing",]$location)))
```

#### Countries
```{r}
country_data <- as.data.frame(table(
    metrics[(metrics$country != "unknown/missing" ) & (metrics$country != ""),]$country
))
colnames(country_data) <- c("Country", "total")
country_data <- arrange(country_data, desc(total))
country_data
```

#### Number of Countries
```{r}
paste0(nrow(country_data), " countries, represented by ", sum(country_data$total), " samples")
```

#### Number of Samples With Isolation Source
```{r}
has_source <- nrow(metrics[metrics$isolation_source != "",])
paste0(has_source," (", has_source / nrow(metrics) * 100, " %)")
```

#### Isolation Sources
```{r}
df <- as.data.frame(table(substr(tolower(
    metrics[metrics$isolation_source != "",]$isolation_source), 1, 50
)))
df[order(-df$Freq),]
```

#### Number of Isolation Sources
```{r}
nrow(as.data.frame(table(tolower(
    metrics[metrics$isolation_source != "",]$isolation_source
))))
```

# Session Info
```{r}
sessionInfo()
```
